---
title: "Assignment 1"
author: "Andrew Tu"
date: "9/27/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
```

## Question 1
Which plane (tailnum) has the worst on-time record?

```{r deptime}
flights%>%
  filter(is.na(dep_time))%>%
  nrow()

```

8255 flights have missing departure time

```{r missing}
which(colSums(is.na(flights))>0)

```
other than dep_time, the columns dep_delay, arr_time, arr_delay, tailnum, and air_time are also missing. Departure delay, arrival time, arrival delay and air time can be NA because a flight was cancelled, tailnum can be missing if it is an unregistered flight or a secret government chartered flight.


## Question 2
Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r timestamp}
flights_2 <- flights %>% 
  mutate(
    dep_time = ((dep_time %/% 100) * 60 + (dep_time %% 100))%%(24*60),
    sched_dep_time = ((sched_dep_time %/% 100) * 60 + (sched_dep_time %% 100))%%(24*60)
    )

flights_2
```

## Question 3
Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()

```{r delay}
flights_3 = flights_2 %>%
  group_by(year,month,day) %>%
  summarise(
    daily_cancelled = sum(is.na(air_time)&is.na(dep_time)&is.na(arr_time)&is.na(arr_delay)),
    mean_arr_delay = mean(arr_delay, na.rm = TRUE),
    mean_dep_delay = mean(dep_delay, na.rm = TRUE)
    )%>%
  ungroup()%>%
  select(c('daily_cancelled','mean_arr_delay','mean_dep_delay'))%>%
  pivot_longer(cols = c('mean_arr_delay','mean_dep_delay'),names_to = 'delay_type', values_to = 'delay')

ggplot(flights_3, aes(x = delay, y = daily_cancelled, color = delay_type))+
  geom_point()+
  labs(x = "Mean delay (minutes per day)", y = "Flights cancelled (per day)", color = 'Type of Delay')+
  ggtitle('Flight delay and flight cancellation at JFK')+
  scale_color_brewer(palette='Set1',labels=c('Arrival', 'Departure'))
  

```

There appears to be a positive correlation between flight delays and flight cancellations of that day, however the correlation does not appear to be linearD