---
title: "hw3"
author: "Andrew Tu"
date: "2022-10-19"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
```

# Q1
Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

```{r q1}
# airports %>%
#   semi_join(flights, c("faa" = "dest")) %>%
#   ggplot(aes(lon, lat)) +
#     borders("state") +
#     geom_point() +
#     coord_quickmap()
airport_loc = airports%>%
  mutate(dest = faa)%>%
  select(dest,lat,lon)

flights_1 = flights%>%
  group_by(dest)%>%
  mutate(avg_delay = (dep_delay+arr_delay)/2)%>%
  summarise(avg_delay = mean(avg_delay,na.rm = TRUE))%>%
  left_join(airport_loc, 'dest')

ggplot(flights_1,aes(lon, lat)) +
  borders("state") +
  geom_point(aes(color = avg_delay)) +
  coord_quickmap()
```

# Q2 
Add the location of the origin and destination (i.e. the lat and lon) to flights.


```{r q2}

airport_ori = airports%>%
  rename(lat_ori = lat)%>%
  rename(lon_ori =  lon)%>%
  select(faa,lon_ori,lat_ori)

airport_dest = airports%>%
  rename(lat_dest = lat)%>%
  rename(lon_dest =  lon)%>%
  select(faa,lat_dest,lon_dest)

flights%>%
  left_join(airport_ori,c("origin" = "faa"))%>%
  left_join(airport_dest,c("dest" = "faa"))
```
# q3
Is there a relationship between the age of a plane and its delays?
```{r q3}

plane_age = planes%>%
  mutate(age = 2013-year)%>%
  select(tailnum,age)

flights_3 = flights%>%
  group_by(tailnum)%>%
  mutate(avg_delay = (dep_delay+arr_delay)/2)%>%
  summarise(avg_delay = mean(avg_delay,na.rm = TRUE))%>%
  left_join(plane_age, 'tailnum')
ggplot(flights_3,aes(age,avg_delay))+
  geom_point()+
  geom_smooth(method = 'lm')

```
There doesn't appear to be any correlation between plane age and aggregate arrival and departure delay. There are a few outliers with high delays in the younger planes, alhough this could entirely be due to there being a greater number of planes under the ae of 25. The linear regression concurs with this finding.