---
title: "hw2"
author: "Andrew Tu"
date: "10/14/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
library(lubridate)
library(magrittr)
```

What time of day should you fly if you want to avoid delays as much as possible? 

```{r time to fly}
flights_dt = flights %>%
  mutate(time = with_tz(as_datetime(time_hour), tzone = "US/Eastern"))

minute(flights_dt$time) <- flights_dt$minute

flights_delay = flights_dt%>%
  group_by(hour)%>%
  summarise(
    mean_arr_delay = mean(arr_delay, na.rm = TRUE),
    mean_dep_delay = mean(dep_delay, na.rm = TRUE))%>%
  pivot_longer(
    cols = c('mean_arr_delay','mean_dep_delay'),
    names_to = 'delay_type', 
    values_to = 'delay')

flights_delay$delay_type <- factor(
  flights_delay$delay_type,
  levels = c("mean_dep_delay","mean_arr_delay"),
  labels = c( "Departure Delay","Arrival Delay"))

```

```{r delays}


ggplot(flights_delay)+
  geom_col(aes(x = hour, y = delay))+
  facet_wrap(~delay_type)+
  ylab('Delay (minutes)')+
  ggtitle('Delays and Scheduled Departure Time')
  

# ggplot(flights_delay)+
#   geom_bar(stat='identity',aes(x = hour, y = mean_dep_delay))
  

```
There appears to be fewer delays early in the morning, with delays peaking out at around 7 to 9 PM.
  

Does this choice depend on anything? Season? Weather? Airport? Airline? Find three patterns (“null results” are ok!). Write your results into Rmarkdown. Include a short introduction that summarizes the three results. Then, have a section for each finding. Support each finding with data summaries and visualizations. Include your code when necessary. This shouldn’t be long, but it might take some time to find the things you want to talk about and lay them out in an orderly way.


```{r season}
flights_season = flights_dt%>%
  mutate(season = quarter(time))%>%
  mutate(season = recode_factor(season, "1" = "Spring", "2" = "Summer", "3" = "Fall", "4" = "Winter"))

ggplot(flights_season, aes(x= season, y = arr_delay))+
  #borrowed from https://stackoverflow.com/questions/48936704/manual-specification-of-draw-quantile-argument-in-violin-plot-call-in-r
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), linetype = "dashed")+
  geom_violin(fill="transparent",draw_quantiles = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=7, color="red", fill="red")+
  ylim(-100,200)+
  ylab("Arrival Delay")
ggplot(flights_season, aes(x= season, y = dep_delay))+
  #borrowed from https://stackoverflow.com/questions/48936704/manual-specification-of-draw-quantile-argument-in-violin-plot-call-in-r
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), linetype = "dashed")+
  geom_violin(fill="transparent",draw_quantiles = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=7, color="red", fill="red")+
  ylim(-30,100)+
  ylab("Departure Delay")
```
There does not appear to be any major differences between the 4 seasons in terms of arrival or departure delays

```{r airport}
flights_airport = flights_dt%>%
  group_by(origin)%>%
  summarise(
    mean_arr_delay = mean(arr_delay, na.rm = TRUE),
    mean_dep_delay = mean(dep_delay, na.rm = TRUE))%>%
  pivot_longer(cols = c("mean_arr_delay","mean_dep_delay"),
               values_to = "delay",
               names_to = "delay_type")
ggplot(flights_airport,aes(x = origin,y = delay, color = delay_type))+
  geom_point(size = 7)+
  scale_color_brewer(palette='Set1',labels=c('Arrival', 'Departure'))+
  guides(color=guide_legend(title="Delay Type"),size = FALSE)+
  ylab('Delay (minutes)')+
  ggtitle('Airport and Average Delay')+
  xlab('Airport')+
  ylim(0,18)
```
Newark is the worst in terms of delays, LaGuardia has less departure delay but greater arrival delay at the destination airport compared to JFK.

```{r weekday}
flight_weekday = flights_dt%>%
  mutate(day_week = wday(time, label = TRUE))%>%
  group_by(day_week)%>%
  summarise(
  mean_arr_delay = mean(arr_delay, na.rm = TRUE),
  mean_dep_delay = mean(dep_delay, na.rm = TRUE))%>%
  pivot_longer(
  cols = c('mean_arr_delay','mean_dep_delay'),
  names_to = 'delay_type', 
  values_to = 'delay')

ggplot(flight_weekday, aes(x = day_week, y = delay, color = delay_type))+
  geom_point(size = 5)+
  scale_color_brewer(palette='Set1',labels=c('Arrival', 'Departure'))+
  guides(color=guide_legend(title="Delay Type"),size = FALSE)+
  ylab('Delay (minutes)')+
  ggtitle('Day of Week and Average Delay')+
  xlab('Day of Week')+
  ylim(-5,25)
```
Earlier and Later in the week there are more delays, with fewer on Tuesday and Wednesdays, and surprisingly fewest delays on Saturday, with planes actually Arriving before schedule.